### 2023-07-13 join.qmd
init

---
title: "Interested in joining the team"
pagetitle: "CarterLab | Join"
toc: false
---

We are always seeking curious, creative, and self-motivated students to enhance our research team and learning environment. The Action, Cognition, & Metascience lab at McMaster University is committed to the training and development of all its team members by promoting a highly collaborative research and training environment, and fostering a passion for lifelong learning and discovery. We are committed to supporting and promoting diversity and equity in our lab. **Everyone---regardless of race, nationality, sex, gender, sexual orientation, age, disability, or religion---is welcome in our lab.**

Updates in progress...