### 2023-07-13 papers/2010-14/2010_patterson_hms.qmd

---
title: "Learner regulated knowledge of results during the acquisition of multiple timing goals"
type: "article"
author: "Patterson JT. & Carter MJ."
year: "2010"
publication: "Human Movement Science"
pdf: ""
doi: "https://doi.org/10.1016/j.humov.2009.12.003"
github: ""
materials: ""
#supplementary: ""
preprint: ""
toc: false
---

## Citation

### APA 7th

> Patterson, J. T., & Carter, M. J. (2010). Learner regulated knowledge of results during the acquisition of multiple timing goals. *Human Movement Science*, 29(2), 214–227. https://doi.org/10.1016/j.humov.2009.12.003

### Bibtex

<pre>
@article{,
  title = {Learner Regulated Knowledge of Results during the Acquisition of Multiple Timing Goals},
  author = {Patterson, Jae T. and Carter, Michael J.},
  date = {2010-04-01},
  journaltitle = {Human Movement Science},
  shortjournal = {Human Movement Science},
  volume = {29},
  number = {2},
  pages = {214--227},
  issn = {0167-9457},
  doi = {10.1016/j.humov.2009.12.003},
  url = {https://www.sciencedirect.com/science/article/pii/S0167945710000035},
  urldate = {2023-07-13},
  langid = {english},
  keywords = {Knowledge of results,Motor learning,Motor skills,Practice,Self-control}
}
</pre>

## Abstract

The purpose of the present experiment was to examine the advantages of a learner controlled KR schedule during the acquisition of three novel sequential timing tasks. The self-regulated group requested KR when necessary during the acquisition period while participants in the yoked condition replicated the KR schedule of a self-regulated counterpart, without the choice. The self-regulated condition demonstrated superior performance in retention and transfer, with a relative KR frequency similar for all three sequences. Similar to Chiviacowsky and Wulf (2002), learners also demonstrated a preference for KR after perceived good trials, independent of defined task difficulty. Thus, the results extend previous research by suggesting a generalized learning strategy by performers acquiring multiple motor task goals.