### 2023-07-13 papers/2015-19/2016_stemarie_jss.qmd

---
title: "Self-controlled learning benefits: exploring contributions of self-efficacy and intrinsic motivation via path analysis"
type: "article"
author: "Ste-Marie DM., Carter MJ., Law B., Vertes K., & Smith V."
year: "2016"
publication: "Journal of Sports Sciences"
pdf: ""
doi: "https://doi.org/10.1080/02640414.2015.1130236"
github: ""
materials: ""
#supplementary: ""
preprint: ""
toc: false
---

## Citation

### APA 7th

> Ste-Marie, D. M., Carter, M. J., Law, B., Vertes, K., & Smith, V. (2016). Self-controlled learning benefits: Exploring contributions of self-efficacy and intrinsic motivation via path analysis. *Journal of Sports Sciences*, 34(17), 1650–1656. https://doi.org/10.1080/02640414.2015.1130236

### Bibtex

<pre>
@article{,
  title = {Self-Controlled Learning Benefits: {{Exploring}} Contributions of Self-Efficacy and Intrinsic Motivation via Path Analysis},
  shorttitle = {Self-Controlled Learning Benefits},
  author = {Ste-Marie, Diane M. and Carter, Michael J. and Law, Barbi and Vertes, Kelly and Smith, Victoria},
  date = {2016-09-01},
  journaltitle = {Journal of Sports Sciences},
  volume = {34},
  number = {17},
  eprint = {26707002},
  eprinttype = {pmid},
  pages = {1650--1656},
  publisher = {{Routledge}},
  issn = {0264-0414},
  doi = {10.1080/02640414.2015.1130236},
  url = {https://doi.org/10.1080/02640414.2015.1130236},
  urldate = {2023-07-13},
  keywords = {knowledge of performance,learner-controlled,Motor skill learning,video feedback}
}
</pre>

## Abstract

Research has shown learning advantages for self-controlled practice contexts relative to yoked (i.e., experimenter-imposed) contexts; yet, explanations for this phenomenon remain relatively untested. We examined, via path analysis, whether self-efficacy and intrinsic motivation are important constructs for explaining self-controlled learning benefits. The path model was created using theory-based and empirically supported relationships to examine causal links between these psychological constructs and physical performance. We hypothesised that self-efficacy and intrinsic motivation would have greater predictive power for learning under self-controlled compared to yoked conditions. Participants learned double-mini trampoline progressions, and measures of physical performance, self-efficacy and intrinsic motivation were collected over two practice days and a delayed retention day. The self-controlled group (*M* = 2.04, *SD* = .98) completed significantly more skill progressions in retention than their yoked counterparts (*M* = 1.3, *SD* = .65). The path model displayed adequate fit, and similar significant path coefficients were found for both groups wherein each variable was predominantly predicted by its preceding time point (e.g., self-efficacy time 1 predicts self-efficacy time 2). Interestingly, the model was not moderated by group; thus, failing to support the hypothesis that self-efficacy and intrinsic motivation have greater predictive power for learning under self-controlled relative to yoked conditions.