### 2023-07-13 papers/2015-19/2017_carter_brainstim.qmd

---
title: "Intentional switches between coordination patterns are faster following anodal-tDCS applied over the supplementary motor area"
type: "article"
author: "Carter MJ., & Maslovat D., & Carlsen AN."
year: "2017"
publication: "Brain Stimulation"
pdf: ""
doi: "https://doi.org/10.1016/j.brs.2016.11.002"
github: ""
materials: ""
#supplementary: ""
preprint: ""
toc: false
---

## Citation

### APA 7th

> Carter, M. J., Maslovat, D., & Carlsen, A. N. (2017). Intentional switches between coordination patterns are faster following anodal-tDCS applied over the supplementary motor area. *Brain Stimulation*, 10(1), 162–164. https://doi.org/10.1016/j.brs.2016.11.002

### Bibtex

<pre>
@article{,
  title = {Intentional Switches between Coordination Patterns Are Faster Following Anodal-{{tDCS}} Applied over the Supplementary Motor Area},
  author = {Carter, Michael J. and Maslovat, Dana and Carlsen, Anthony N.},
  date = {2017-01-01},
  journaltitle = {Brain Stimulation},
  shortjournal = {Brain Stimulation},
  volume = {10},
  number = {1},
  pages = {162--164},
  issn = {1935-861X},
  doi = {10.1016/j.brs.2016.11.002},
  url = {https://www.sciencedirect.com/science/article/pii/S1935861X1630300X},
  urldate = {2023-07-13},
  langid = {english}
}
</pre>