### 2023-07-13 team/senior/mikayla-lalli.qmd

---
title: "Mikayla Lalli"
subtitle: "M.Sc. Candidate"
image: mikayla-lalli.jpg
sortby: 2
started: "September 2022"
about:
  id: hero-heading
  template: solana
  links:
    - icon: github
      text: github
      href: https://www.github.com/mikaylalalli
---

:::{#hero-heading}

I joined the lab in September 2022 after completing my undergraduate studies at McMaster University in the School of Interdisciplinary Science. I am interested in motor cognition, specifically understanding the underlying cognitive mechanisms of motor interactions between multiple agents. My research focuses on decision making in collaborative human-human interactions. In my free time I enjoy playing the drums and teaching the lab about Harry Styles.

## Education

**McMaster University** | Hamilton, ON, Canada  
M.Sc in Kinesiology, Department of Kinesiology  
2022--Present

**McMaster University** | Hamilton, ON, Canada  
B.Sc. (Honours) in Life Sciences (Minor in Psychology), School of Interdisciplinary Science  
2018--22

:::