### 2023-07-13 team/senior/mike-carter.qmd

---
title: "Michael Carter"
subtitle: "Principal Investigator"
image: mike-carter.jpg
sortby: 1
started: "July 2017"
#resources:
  #- "carter_cv.pdf"
about:
  id: hero-heading
  template: solana
  links:
    #- icon: file-earmark-text
    #  text: vitae
    #  href: ./team/senior/carter_cv.pdf
    - icon: mortarboard
      text: scholar
      href: https://scholar.google.com/citations?user=9FY5gBwAAAAJ
    - icon: github
      text: github
      href: https://www.github.com/carterjmike
    - icon: twitter
      text: twitter
      href: https://www.twitter.com/carterjmike
---

:::{#hero-heading}

I am broadly interested in how the brain controls and learns skilled actions, as well as how we can improve the replicability and reproducibility of kinesiology and sport science research. This work relies on an interdisciplinary blend of theory, behaviour experimentation, and computational approaches from psychology, neuroscience, engineering, computer science, and statistics.

I completed my doctoral training under the supervision of [Dr. Diane Ste-Marie](https://www.uottawa.ca/faculty-health-sciences/human-kinetics/our-professors/diane-ste-marie) (Emeritus Professor) at the University of Ottawa where I studied the cognitive processes underlying self-controlled practice. During this time, I also worked closely with [Dr. Tony Carlsen](https://neuromotor.ca/) on projects related to bimanual coordination and motor preparation. I then completed a one-year postdoctoral fellowship in the Centre for Neuroscience Studies at Queen's University under the mentorship of [Dr. Randy Flanagan](https://www.flanganlab.com) and [Dr. Jason Gallivan](https://www.gallivanmaplab.com).

I am currently an Assistant Professor in the Department of Kinesiology (joined in 2017) at McMaster University, where I direct the Action, Cognition, and Metascience Lab. I teach undergraduate and graduate courses in the areas of motor control and learning, neurorehabilitation and robotics, statistics, and scientific computing. I currently serve as a Section Editor (Sensorimotor Control) for the journal Communications in Kinesiology, which is the flagship journal for the [Society of Transparency, Openness, and Replication in Kinesiology](http://www.storkinesiology.org/) (STORK). I also currently serve as the Chair of STORK.

## Education

**Queen's University** | Kingston, ON, Canada  
Postdoctoral Fellow, Centre for Neuroscience Studies  
2016--17

**University of Ottawa** | Ottawa, ON, Canada  
Ph.D. in Human Kinetics, School of Human Kinetics  
2011--16

**Brock University** | St. Catharines, ON, Canada  
M.Sc. in Kinesiology, Department of Kinesiology
2008--10

**Brock University** | St. Catharines, ON, Canada  
Bachelor of Physical Education (Honours), Department of Kinesiology
2004--08

:::