### 2023-08-29 team/senior/nour-alafif.qmd

---
title: "Nour Al Afif"
subtitle: "M.Sc. Candidate"
image: nour-alafif.jpg
sortby: 4
started: "September 2022"
about:
  id: hero-heading
  template: solana
  links:
    - icon: github
      text: github
      href: https://www.github.com/NourAlAfif19
---

:::{#hero-heading}

I joined the lab in September 2022 after completing my undergraduate studies at McMaster University in the Department of Psychology, Neuroscience, and Behaviour. My research interests include sensorimotor interactions (e.g., human-agent and human-human) and how different partners and forms of interaction influence motor learning. Other things I contribute to the lab include attempting to help with code and also making everyone feel better about themselves by being impressively bad at bowling.

## Education

**McMaster University** | Hamilton, ON, Canada  
M.Sc in Kinesiology, Department of Kinesiology  
2022--Present

**McMaster University** | Hamilton, ON, Canada  
B.Sc. (Honours) in Psychology, Neuroscience, & Behaviour (Minor in Biology), Department of Psychology, Neuroscience, & Behaviour  
2018--22

:::