### 2023-08-29 team/senior/daniel-deletsu.qmd
init

---
title: "Daniel Deletsu"
subtitle: "M.Sc. Student"
image: daniel-deletsu.jpg
sortby: 5
started: "September 2023"
about:
  id: hero-heading
  template: solana
---

:::{#hero-heading}

I joined the lab in September 2023 after completing my undergraduate studies at Queen’s University where I majored in Life Sciences. I am interested in exploring the how sensory, motor, and cognitive processes are affected in healthy aging and in neurological disorders such as Alzheimer's disease. Outside of the lab, you are most likely to find me in a gym or on a basketball court.

## Education

**McMaster University** | Hamilton, ON, Canada  
M.Sc in Kinesiology, Department of Kinesiology  
2023--Present

**Queen's University** | Kingston, ON, Canada  
B.Sc. (Honours) in Life Sciences, Faculty of Health Sciences  
2019--23

:::