### 2023-08-29 team/senior/jack-tang.qmd

---
title: "Jiaqiao (Jack) Tang"
subtitle: "Ph.D. Student"
image: jack-tang.jpg
sortby: 2
started: "September 2023"
about:
  id: hero-heading
  template: solana
---

:::{#hero-heading}

I joined the lab in September 2023 after completing my Master of Science in Physical Therapy at the University of Toronto. Prior to that, I completed an undergraduate degree here at McMaster University where I earned my Bachelor of Science in Kinesiology. My research interests include the impact of exercise and/or physical activity on optimal aging, in particular cognition and mobility. My extracurricular interests include rock climbing and finding new places to eat!

## Education

**McMaster University** | Hamilton, ON, Canada  
Ph.D. in Kinesiology, Department of Kinesiology  
2022--Present

**University of Toronto** | Toronto, ON, Canada
M.Sc. in Physical Therapy, Department of Physical Therapy
2016--18

**McMaster University** | Hamilton, ON, Canada  
B.Sc. (Honours) in Kinesiology, Department of Kinesiology  
2011--15

:::