### 2023-09-22 team/junior/ola-schwarzenberg.qmd

---
title: "Ola Schwarzenberg"
subtitle: "Undergraduate thesis student"
image: ola-schwarzenberg.jpg
sortby: 3
started: "September 2023"
about:
  id: hero-heading
  template: solana
---

:::{#hero-heading}

I joined the lab in September 2023 as an undergraduate thesis student and I am in the final year of my undergraduate degree in Life Sciences (Sensory Motor Systems specialization) with a minor in English and Cultural Studies. For my thesis work, I will be investigating cooperative decision-making. In my free time, I like to read, run, and play the piano.

:::