### 2023-09-22 team/junior/mithun-manivannan.qmd
add new member

---
title: "Mithun Manivannan"
subtitle: "Undergraduate thesis student"
image: mithun-manivannan.jpg
sortby: 4
started: "September 2023"
about:
  id: hero-heading
  template: solana
---

:::{#hero-heading}

I joined the lab in September 2023 as an undergraduate thesis student and I am in my fourth year pursuing both a Honours Kinesiology and Mathematical Sciences Degree with a minor in Psychology. For my thesis work, I will be investigating sensorimotor human-human interactions and rehabilitation. Outside the lab, I enjoy playing the guitar, reading and staying active with friends.

:::