### 2023-09-23 team/junior/enuri-dissanayake.qmd
fix typo

---
title: "Enuri Dissanayake"
subtitle: "Undergraduate practicum student"
image: enuri-dissanayake.jpg
sortby: 5
started: "September 2023"
about:
  id: hero-heading
  template: solana
---

:::{#hero-heading}

I joined the lab in September 2023 as a practicum student, and I am currently in my third year of my undergraduate degree in Life Sciences (Sensory Motor Systems). My research interests include motor learning and collaborative decision-making. Outside of the lab, I enjoy dragonboating, working out with friends, and finding new music!

:::