### 2023-09-27 team/junior/jacqueline-zheng.qmd
add new members to team page

---
title: "Jacqueline Zheng"
subtitle: "Undergraduate practicum student"
image: jacqueline-zheng.jpg
sortby: 6
started: "September 2023"
about:
  id: hero-heading
  template: solana
---

:::{#hero-heading}

I joined the lab in September 2023 as a research practicum student and I am in my third year of the Honours Neuroscience program. My research interests include sensorimotor interactions and motor learning metascience. Outside of lab, I am a cat parent who loves food hunting in Hamilton and the GTA.

:::