### 2024-01-21 papers.qmd
change dir date

---
title: "Papers"
pagetitle: "CarterLab | Papers"
toc: true
toc-title: "Jump to..."
listing:
  - id: preprints
    contents: "papers/preprints/*.qmd"
    template: papers.ejs
    sort:
      - "year desc"
      - "author"
  - id: papers2020
    contents: "papers/2020-24/*.qmd"
    template: papers.ejs
    sort:
      - "year desc"
      - "author"
  - id: papers2015
    contents: "papers/2015-19/*.qmd"
    template: papers.ejs
    sort:
      - "year desc"
      - "author"
  - id: papers2010
    contents: "papers/2010-14/*.qmd"
    template: papers.ejs
    sort:
      - "year desc"
      - "author"
---

## Preprints

:::{#preprints}
:::

## 2020--24

:::{#papers2020}
:::

## 2015--19

:::{#papers2015}
:::

## 2010--14

:::{#papers2010}
:::