### 2024-02-06 team/junior/francesca-caputo.qmd
add new members

---
title: "Francesca Caputo"
subtitle: "Undergraduate practicum student"
image: francesca-caputo.png
sortby: 6
started: "January 2024"
about:
  id: hero-heading
  template: solana
---

:::{#hero-heading}

I joined the lab in January 2024 as a practicum student. I am currently in the third year of my undergraduate degree in the Biology Research Specialization. My research interests include motor learning and the dynamics of neural circuits. Outside of the lab, I enjoy testing out new recipes, going on adventures with my friends, and playing the guitar!

:::