### 2024-02-06 team/junior/jil-soni.qmd

---
title: "Jil Soni"
subtitle: "Undergraduate practicum student"
image: jil-soni.png
sortby: 6
started: "January 2024"
about:
  id: hero-heading
  template: solana
---

:::{#hero-heading}

I joined the lab in January 2024 as a research practicum student. I am in my third year of Life Sciences (Sensory Motor Systems specialization) with a minor in Statistics. My research interests include sensorimotor interactions and cooperative decision-making. Outside of the lab, I enjoy reading a good book, painting, and going on hikes!

:::