### 2024-04-12 team/junior/michael-croteau.qmd

---
title: "Michael Croteau"
subtitle: "Undergraduate research assistant"
image: michael-croteau.jpg
sortby: 6
started: "January 2024"
about:
  id: hero-heading
  template: solana
---

:::{#hero-heading}

I joined the lab in January 2024 as a practicum student. I am in my third year as an undergraduate neuroscience student. My research interests include investigating somatosensory and motor systems, and I also enjoy learning about computer science and organic chemistry on the side. In my spare time I like to rock climb, work out, read (mostly sci-fi and fantasy), and play video games and sports.

:::