### 2024-04-12 team.qmd
update team page

---
title: "Meet the team"
toc: false
listing:
  - id: senior
    contents: "team/senior/*.qmd"
    type: grid
    image-height: 225px
    grid-columns: 3
    fields: 
      - image
      - title
      - subtitle
    sort: sortby
    field-display-names: 
      started: "Started"
  - id: junior
    contents: "team/junior/*.qmd"
    type: grid
    image-height: 225px
    grid-columns: 3
    fields: 
      - image
      - title
      - subtitle
    sort: sortby
    field-display-names: 
      started: "Started"
---

## Senior members

:::{#senior}
:::

## Junior members

:::{#junior}
:::

## Collaborators

The lab maintains active and extensive research collaborations with:  
[Joshua Cashaback](https://joshcashaback.weebly.com/) | Department of Biomedical Engineering, University of Delaware  
[Brad Mckay](https://bradmckay-motor-meta.netlify.app/) | Milwaukee Brewers

Some of our other collaborators include:  
[Keith Lohse](https://sites.google.com/site/lohsekr/) | Department of Neurology, Washington University School of Medicine in Saint Louis  
[Matthew Miller](http://www.education.auburn.edu/initiatives/performance-exercise-psychophysiology-lab-pep-matt-miller-ph-d/) | School of Kinesiology, Auburn University  
[Mariane Bacelar](https://www.boisestate.edu/kinesiology/faculty-and-staff-mariane-bacelar-phd/) | Department of Kinesiology, Boise State University

## Alumni

Former lab members and what they did after leaving the lab or what they are currently doing (if known).

### Post-doctoral fellows

**Brad McKay** (2021-23); *Senior Strategist of Training Design, Milwaukee Brewers*

### Graduate students

**Laura St. Germain**, Ph.D. Kinesiology (2018-23); *Research and Data Specialist at Sheridan College*  
**Mirette Mounir**, M.Sc. Kinesiology (2019-21); *Dentistry at Western University*  
**Rajbir Sihu**, M.Sc. Kinesiology (2019-21); *Dentistry at the University of Toronto*

### Undergraduate students

#### Honours thesis

Lidia Barbera (2023-24); Hibaa Hasan (2023-24); Ola Schwarzenberg (2023-24); Mithun Manivannan (2023-24); Chitrini Tandon (2021-23); Alyssa Skouros (2022-23); Kiana Masoudi (2022-23); Raika Bourmand (2021-22); Armin Sariaslani (2021-22); Katharine (Katie) Douglas (2021-22); Thinzar Soe (2021-22); Julianna Marfisi (2020-21); Matthew Tobis (2020-21); Andrew Poskus (2020-21); Sherry Feldman (2019-20); Olena Leshchyshen (2019-20); Noura Balbaa (2019-20)

#### Research practicum / project

Jacqueline Zheng (2023-24); Abby Girouard (2023-24); Michael Croteau (2024); Aditya Trivedi (2024); Navishka Brahmbhatt (2024); Francesca Caputo (2024); Jil Soni (2024); Enuri Dissanayake (2023); Jeswende Seedu (2022-23); Vivian Li (2022); Caroline Jarvi (2022); Gurlal Gill (2021); Anita Chaseendran (2021); Gianna Jeyarajan (2021); Mackenzie (Kenzie) Bell (2021); Ashton Yuen (2021); Charlie Violin (2021); Ammaar Jan (2020); Faran Chaudhry (2020); Yu (Sarah) Fu (2020); Andrew Poskus (2019); Maya El-Zahed (2019); Grant Yang (2019); Olena Leshchyshen (2019); Rita Waseem (2019); Christian Kleiser (2018); Christopher Li (2018); Faryal Zahir (2018)

#### Research assistants

Elizabeth Mitchell (2020-23); Celeste de Faveri (2021-23); Lidia Barbera (2022-23); Hibaa Hasan (2022-23); Faith Adams (2022); Chitrini Tandon (2021-22); Kristen Arnold (2021-22); Alison Williams (2018-20); Andrew Poskus (2019); Kevin Gilotra (2019); Ashton Yuen (2019); Mackenzie (Kenzie) Bell (2019-21); Olena Leshchyshen (2018-19); Yu (Sarah) Fu (2018-19)