### 2024-05-17 team/junior/eleanore-wu.qmd
add new member

---
title: "Eleanore Wu"
subtitle: "Undergraduate research assistant"
image: eleanore-wu.png
sortby: 6
started: "May 2024"
about:
  id: hero-heading
  template: solana
---

:::{#hero-heading}

I joined the lab as an undergraduate research assistant in May 2024. My research interests include human-agent and human-human interactions, as well as investigating how these factors could affect motor learning. Outside of the lab, I enjoy playing badminton and volleyball, baking, and drinking bubble tea.

:::