### 2024-05-31 team/junior/eshaa-tariq.qmd
update team page

---
title: "Eshaa Tariq"
subtitle: "Undergraduate research assistant"
image: eshaa-tariq.jpg
sortby: 6
started: "May 2024"
about:
  id: hero-heading
  template: solana
---

:::{#hero-heading}

I joined the lab in May 2024 as an undergraduate research assistant. I am currently enrolled in the Honours Kinesiology program at McMaster. My research interests include proprioception and how it particularly relates to sensorimotor interactions. Outside of the lab, I enjoy playing volleyball, reading fictional novels, and watching true crime and horror shows.

:::