### 2025-01-17 team.qmd

---
title: "Meet the team"
toc: false
---

<style>
.team-member {
  display: flex;
  align-items: center;
  padding: 20px; /* Add padding around each team member */
  border: 1px solid #ddd; /* Optional: Add a border for better visual separation */
  margin-bottom: 20px; /* Add space between each team member */
}

.team-img {
  margin-right: 20px;
  width: 175px; /* Set the desired width */
  height: 175px; /* Set the desired height */
  border-radius: 50%; /* Make the image rounded */
  object-fit: cover; /* Ensure the image covers the area */
}

.team-info {
  display: flex;
  flex-direction: column;
}
</style>

<div class="team-member">
  <img src="images/mike-carter.jpg" alt="images/mike-carter.jpg" class="team-img">
  <div class="team-info">
  <h4> Mike Carter, Ph.D. </h4>
  <p> Assistant Professor | Principal Investigator </p>
  <p> cartem11 [at] mcmaster [dot] ca </p>
  </div>
</div>

<div class="team-member">
  <img src="images/jack-tang.jpg" alt="images/jack-tang.jpg" class="team-img">
  <div class="team-info">
  <h4> Jiaqiao Tang, M.Sc.PT </h4>
  <p> Ph.D. Student </p>
  <p> tangj24 [at] mcmaster [dot] ca</p>
  </div>
</div>

<div class="team-member">
  <img src="images/mikayla-lalli.jpg" alt="images/mikayla-lalli.jpg" class="team-img">
  <div class="team-info">
  <h4> Mikayla Lalli, M.Sc. </h4>
  <p> Ph.D. Student </p>
  <p> lallim1 [at] mcmaster [dot] ca </p>
  </div>
</div>

<div class="team-member">
  <img src="images/nour-al-afif.jpg" alt="images/nour-al-afif.jpg" class="team-img">
  <div class="team-info">
  <h4> Nour Al Afif, M.Sc. </h4>
  <p> Ph.D. Student </p>
  <p> alafin1 [at] mcmaster [dot] ca </p>
  </div>
</div>

<div class="team-member">
  <img src="images/daniel-deletsu.jpg" alt="images/nour-alfif.jpg" class="team-img">
  <div class="team-info">
  <h4> Daniel Deletsu, B.Sc. </h4>
  <p> M.Sc. Student </p>
  <p> deletsud [at] mcmaster [dot] ca </p>
  </div>
</div>

## Collaborators

The lab maintains active and extensive research collaborations with:  
[Joshua Cashaback](https://joshcashaback.weebly.com/) | Department of Biomedical Engineering, University of Delaware  
[Brad Mckay](https://bradmckay-motor-meta.netlify.app/) | Milwaukee Brewers

Some of our other collaborators include:  
[Keith Lohse](https://sites.google.com/site/lohsekr/) | Department of Neurology, Washington University School of Medicine in Saint Louis  
[Matthew Miller](http://www.education.auburn.edu/initiatives/performance-exercise-psychophysiology-lab-pep-matt-miller-ph-d/) | School of Kinesiology, Auburn University  
[Mariane Bacelar](https://www.boisestate.edu/kinesiology/faculty-and-staff-mariane-bacelar-phd/) | Department of Kinesiology, Boise State University

## Alumni

Former lab members and what they did after leaving the lab or what they are currently doing (if known).

### Post-doctoral fellows

**Brad McKay** (2021-23); *Senior Strategist of Training Design, Milwaukee Brewers*

### Graduate students

**Mikayla Lalli**, M.Sc. Kinesiology (2022-24); *Ph.D. student in the lab*.  
**Nour Al Afif**, M.Sc. Kinesiology (2022-24); *Ph.D. student in the lab*  
**Laura St. Germain**, Ph.D. Kinesiology (2018-23); *Research and Data Specialist at Sheridan College*  
**Mirette Mounir**, M.Sc. Kinesiology (2019-21); *Dentistry at Western University*  
**Rajbir Sihu**, M.Sc. Kinesiology (2019-21); *Dentistry at the University of Toronto*

### Undergraduate students

#### Honours thesis

Lidia Barbera (2023-24); Hibaa Hasan (2023-24); Ola Schwarzenberg (2023-24); Mithun Manivannan (2023-24); Chitrini Tandon (2021-23); Alyssa Skouros (2022-23); Kiana Masoudi (2022-23); Raika Bourmand (2021-22); Armin Sariaslani (2021-22); Katharine (Katie) Douglas (2021-22); Thinzar Soe (2021-22); Julianna Marfisi (2020-21); Matthew Tobis (2020-21); Andrew Poskus (2020-21); Sherry Feldman (2019-20); Olena Leshchyshen (2019-20); Noura Balbaa (2019-20)

#### Research practicum / project

Jacqueline Zheng (2023-24); Abby Girouard (2023-24); Michael Croteau (2024); Aditya Trivedi (2024); Navishka Brahmbhatt (2024); Francesca Caputo (2024); Jil Soni (2024); Enuri Dissanayake (2023); Jeswende Seedu (2022-23); Vivian Li (2022); Caroline Jarvi (2022); Gurlal Gill (2021); Anita Chaseendran (2021); Gianna Jeyarajan (2021); Mackenzie (Kenzie) Bell (2021); Ashton Yuen (2021); Charlie Violin (2021); Ammaar Jan (2020); Faran Chaudhry (2020); Yu (Sarah) Fu (2020); Andrew Poskus (2019); Maya El-Zahed (2019); Grant Yang (2019); Olena Leshchyshen (2019); Rita Waseem (2019); Christian Kleiser (2018); Christopher Li (2018); Faryal Zahir (2018)

#### Research assistants

Elizabeth Mitchell (2020-23); Celeste de Faveri (2021-23); Lidia Barbera (2022-23); Hibaa Hasan (2022-23); Faith Adams (2022); Chitrini Tandon (2021-22); Kristen Arnold (2021-22); Alison Williams (2018-20); Andrew Poskus (2019); Kevin Gilotra (2019); Ashton Yuen (2019); Mackenzie (Kenzie) Bell (2019-21); Olena Leshchyshen (2018-19); Yu (Sarah) Fu (2018-19)