### 2025-04-22 index.qmd
update ugrad positions

---
title: ""
pagetitle: "CarterLab | Home"
toc: false
---

![](images/iwc-front.jpg)

The **Action, Cognition, and Metascience Lab** is an interdisciplinary research team interested in how the human brain controls and learns skilled actions in healthy and clinical populations. We also conduct metascience research to gain insights into research practices within kinesiology and related areas. The lab is located in the Ivor Wynne Centre, which is home to the [Department of Kinesiology](https://www.science.mcmaster.ca/kinesiology/){target="_blank"} within the [Faculty of Science](https://www.science.mcmaster.ca/){target="_blank"} at [McMaster University](https://www.mcmaster.ca/){target="_blank"}.

#### Recent lab news ---------

<font color="#d08770">[April 22 2025].</font> **Announcement**. The application period for practicum, project, or thesis positions in the lab for the 2025-26 academic year is now closed.

<font color="#d08770">[March 03 2025].</font> **New Preprint**. Lalli, M., Al Afif, N., Tang, J., Hasan, H., Dissanayake, E., Sussman, V., Lokesh, R., Rathwell, S., Cashaback, J.G.A., & Carter, M.J. Preferences for costly cooperation are highly individualized. *PsyAr$\chi$iv*. [https://doi.org/10.31234/osf.io/3qaeu_v3](https://doi.org/10.31234/osf.io/3qaeu_v3){target="_blank"}