### 2025-05-14 teaching.qmd

---
title: ""
pagetitle: "CarterLab | Teaching"
toc: false
---

```{r}
#| echo: false

library(markdown)
library(tinytable)

tibble::tibble(
    `FALL 2025` = c(
        "Motor Control and Learning (KINESIOL 2MC3)...*Coming soon*",
        "Neurorehabilitation and Robotics (KINESIOL 4TT3)...*Coming soon*",
        "Scientific Computing for Reproducible Science (KINESIOL 736)...*Coming soon*"
        )
) |> 
    tt(width = 1) |> 
    group_tt(i = list(
        "Undergraduate courses" = 1,
        "Graduate courses" = 3
    )) |> 
    format_tt(j = 1, markdown = TRUE) |> 
    style_tt(i = 0, background = "#0077bb", color = "white") |> 
    style_tt(i = c(1,4), background = "#eceff4", bold = TRUE)
```