### 2025-05-14 team.qmd

---
title: "Meet the team"
toc: false
---

<style>
.team-member {
  display: flex;
  align-items: center;
  padding: 20px; 
  border: 1px solid #ddd;
  margin-bottom: 20px;
}

.team-img {
  margin-right: 20px;
  width: 175px;
  height: 175px;
  border-radius: 50%;
  object-fit: cover;
}

.team-info {
  display: flex;
  flex-direction: column;
}
</style>

<div class="team-member">
  <img src="images/mike-carter.jpg" alt="images/mike-carter.jpg" class="team-img">
  <div class="team-info">
  <h4> Mike Carter, Ph.D. </h4>
  <p> Assistant Professor | Principal Investigator </p>
  <p> cartem11 [at] mcmaster [dot] ca </p>
  </div>
</div>

<div class="team-member">
  <img src="images/jack-tang.jpg" alt="images/jack-tang.jpg" class="team-img">
  <div class="team-info">
  <h4> Jiaqiao Tang, M.Sc.PT </h4>
  <p> Ph.D. Student </p>
  <p> tangj24 [at] mcmaster [dot] ca</p>
  </div>
</div>

<div class="team-member">
  <img src="images/mikayla-lalli.jpg" alt="images/mikayla-lalli.jpg" class="team-img">
  <div class="team-info">
  <h4> Mikayla Lalli, M.Sc. </h4>
  <p> Ph.D. Student </p>
  <p> lallim1 [at] mcmaster [dot] ca </p>
  </div>
</div>

<div class="team-member">
  <img src="images/nour-al-afif.jpg" alt="images/nour-al-afif.jpg" class="team-img">
  <div class="team-info">
  <h4> Nour Al Afif, M.Sc. </h4>
  <p> Ph.D. Student </p>
  <p> alafin1 [at] mcmaster [dot] ca </p>
  </div>
</div>

<div class="team-member">
  <img src="images/daniel-deletsu.jpg" alt="images/daniel-deletsu.jpg" class="team-img">
  <div class="team-info">
  <h4> Daniel Deletsu, B.Sc. </h4>
  <p> M.Sc. Candidate </p>
  <p> deletsud [at] mcmaster [dot] ca </p>
  </div>
</div>

<div class="team-member">
  <img src="images/michael-croteau.jpg" alt="images/michael-croteau.jpg" class="team-img">
  <div class="team-info">
  <h4> Michael Croteau </h4>
  <p> M.Sc. Student </p>
  <p> croteaum [at] mcmaster [dot] ca </p>
  </div>
</div>

<div class="team-member">
  <img src="images/lab-logo.png" alt="images/" class="team-img">
  <div class="team-info">
  <h4> Maria Alex </h4>
  <p> Undergraduate Research Assistant </p>
  <p> McMaster Institute for Research on Aging 2025 Summer Research Fellowship Winner </p>
  </div>
</div>

<div class="team-member">
  <img src="images/headshot_duck-bryan.jpg" alt="images/headshot_duck-bryan.jpg" class="team-img">
  <div class="team-info">
  <h4> Duck Bryan </h4>
  <p> Lab duck </p>
  </div>
</div>

<div class="team-member">
  <img src="images/headshot_duck-joy.jpg" alt="images/headshot_duck-joy.jpg" class="team-img">
  <div class="team-info">
  <h4> Duck Joy </h4>
  <p> Lab duck </p>
  </div>
</div>

<div class="team-member">
  <img src="images/headshot_duck-norris.jpg" alt="images/headshot_duck-norris.jpg" class="team-img">
  <div class="team-info">
  <h4> Duck Norris </h4>
  <p> Lab duck </p>
  </div>
</div>

<div class="team-member">
  <img src="images/headshot_ducky-parton.jpg" alt="images/headshot_ducky-parton.jpg" class="team-img">
  <div class="team-info">
  <h4> Ducky Parton </h4>
  <p> Lab duck </p>
  </div>
</div>

<div class="team-member">
  <img src="images/headshot_lil-quack.jpg" alt="images/headshot_lil-quack.jpg" class="team-img">
  <div class="team-info">
  <h4> Lil Quack </h4>
  <p> Lab duck </p>
  </div>
</div>

## Collaborators

The lab maintains active and extensive research collaborations with:  
[Joshua Cashaback](https://joshcashaback.weebly.com/) | Department of Biomedical Engineering, University of Delaware  
[Brad Mckay](https://bradmckay-motor-meta.netlify.app/) | Milwaukee Brewers  
[Jonathan Cannon](https://metre.mcmaster.ca/) | Department of Psychology, Neuroscience, and Behaviour, McMaster University

Some of our other collaborators include:  
[Keith Lohse](https://sites.google.com/site/lohsekr/) | Department of Neurology, Washington University School of Medicine in Saint Louis  
[Matthew Miller](http://www.education.auburn.edu/initiatives/performance-exercise-psychophysiology-lab-pep-matt-miller-ph-d/) | School of Kinesiology, Auburn University  
[Mariane Bacelar](https://www.boisestate.edu/kinesiology/faculty-and-staff-mariane-bacelar-phd/) | Department of Kinesiology, Boise State University

## Alumni

Former lab members and what they did after leaving the lab or what they are currently doing (if known).

### Post-doctoral fellows

**Brad McKay** (2021-23); *Senior Strategist of Training Design, Milwaukee Brewers*

### Graduate students

**Mikayla Lalli**, M.Sc. Kinesiology (2022-24); *Ph.D. student in the lab*.  
**Nour Al Afif**, M.Sc. Kinesiology (2022-24); *Ph.D. student in the lab*  
**Laura St. Germain**, Ph.D. Kinesiology (2018-23); *Research and Data Specialist at Sheridan College*  
**Mirette Mounir**, M.Sc. Kinesiology (2019-21); *Dentistry at Western University*  
**Rajbir Sihu**, M.Sc. Kinesiology (2019-21); *Dentistry at the University of Toronto*

### Undergraduate students

#### Honours thesis

Shiuli Khanna (2024-25); Michael Croteau (2024-25); Enuri Dissanayake (2024-25); Nicole Su (2024-25); Courtney Nguyen (2024-25); Leo Ducharm (2024-25); Andrea Albert Pravin (2024-25); Jacqueline Zhang (2024-25); Lidia Barbera (2023-24); Hibaa Hasan (2023-24); Ola Schwarzenberg (2023-24); Mithun Manivannan (2023-24); Chitrini Tandon (2021-23); Alyssa Skouros (2022-23); Kiana Masoudi (2022-23); Raika Bourmand (2021-22); Armin Sariaslani (2021-22); Katharine (Katie) Douglas (2021-22); Thinzar Soe (2021-22); Julianna Marfisi (2020-21); Matthew Tobis (2020-21); Andrew Poskus (2020-21); Sherry Feldman (2019-20); Olena Leshchyshen (2019-20); Noura Balbaa (2019-20)

#### Research practicum / project

Vida Sussman (2024-25); Tara Esbati (2024-25); Shiv Peshwa (2025); Linsea Martin (2024-25); Faris Almoni (2025); Gaile Cheng (2025); Aashna Gujral (2024); Dami Alabi Benson (2024); Michael Croteau (2024); Francesca Caputo (2024); Jil Soni (2024); Navishka Brahmbhatt (2024); Aditya Trivedi (2024); Jacqueline Zheng (2023-24); Abby Girouard (2023-24); Michael Croteau (2024); Aditya Trivedi (2024); Navishka Brahmbhatt (2024); Francesca Caputo (2024); Jil Soni (2024); Enuri Dissanayake (2023); Jeswende Seedu (2022-23); Vivian Li (2022); Caroline Jarvi (2022); Gurlal Gill (2021); Anita Chaseendran (2021); Gianna Jeyarajan (2021); Mackenzie (Kenzie) Bell (2021); Ashton Yuen (2021); Charlie Violin (2021); Ammaar Jan (2020); Faran Chaudhry (2020); Yu (Sarah) Fu (2020); Andrew Poskus (2019); Maya El-Zahed (2019); Grant Yang (2019); Olena Leshchyshen (2019); Rita Waseem (2019); Christian Kleiser (2018); Christopher Li (2018); Faryal Zahir (2018)

#### Research assistants

Eshaa Tariq (2024-25); Eleanore Wu (2024-25); Shiv Peshwa (2024); Dana Rugole (2025); Vida Sussman (2022-24); Elizabeth Mitchell (2020-23); Celeste de Faveri (2021-23); Lidia Barbera (2022-23); Hibaa Hasan (2022-23); Faith Adams (2022); Chitrini Tandon (2021-22); Kristen Arnold (2021-22); Alison Williams (2018-20); Andrew Poskus (2019); Kevin Gilotra (2019); Ashton Yuen (2019); Mackenzie (Kenzie) Bell (2019-21); Olena Leshchyshen (2018-19); Yu (Sarah) Fu (2018-19)