### 2025-07-29 index.qmd
fix author orderings

---
title: ""
pagetitle: "CarterLab | Home"
toc: false
---

![](images/iwc-front.jpg)

The **Action, Cognition, and Metascience Lab** is an interdisciplinary research team interested in how the human brain controls and learns skilled actions in healthy and clinical populations. We also conduct metascience research to gain insights into research practices within kinesiology and related areas. The lab is located in the Ivor Wynne Centre, which is home to the [Department of Kinesiology](https://www.science.mcmaster.ca/kinesiology/){target="_blank"} within the [Faculty of Science](https://www.science.mcmaster.ca/){target="_blank"} at [McMaster University](https://www.mcmaster.ca/){target="_blank"}.

#### Recent lab news ---------

<font color="#d08770">[June 13 2024].</font> **Paper accepted**. Calalo JA, Ngo T, Sullivan S, Strand K, Buggeln JH, Lokesh R, Roth A, Carter MJ, Kurtzer IL, & Cashaback JGA. (Accepted). Online movements reflect ongoing deliberation. *Journal of Neuroscience*. [https://doi.org/10.1523/JNEUROSCI.1913-24.2025](https://doi.org/10.1523/JNEUROSCI.1913-24.2025){target="_blank"}

<font color="#d08770">[May 14 2024].</font> **Paper accepted**. Sullivan SR, Lokesh R, Calalo J, Ngo TT, Buggeln JH, Roth AR, Peters CS, Kurtzer IL, Carter MJ, & Cashaback JGA. (In Press). Indecision under time pressure arises from suboptimal switching behaviour. *Journal of Neurophysiology*. [https://doi.org/10.1152/jn.00563.2024](https://doi.org/10.1152/jn.00563.2024){target="_blank"}

<font color="#d08770">[April 22 2025].</font> **Announcement**. The application period for practicum, project, or thesis positions in the lab for the 2025-26 academic year is now closed.

<font color="#d08770">[March 03 2025].</font> **New Preprint**. Lalli M, Al Afif N, Tang J, Hasan H, Dissanayake E, Sussman V, Lokesh R, Rathwell S, Cashaback JGA, & Carter MJ. Preferences for costly cooperation are highly individualized. *PsyAr$\chi$iv*. [https://doi.org/10.31234/osf.io/3qaeu_v3](https://doi.org/10.31234/osf.io/3qaeu_v3){target="_blank"}