### 2025-08-01 index.qmd
add new preprint info

---
title: ""
pagetitle: "CarterLab | Home"
toc: false
---

![](images/iwc-front.jpg)

The **Action, Cognition, and Metascience Lab** is an interdisciplinary research team interested in how the human brain controls and learns skilled actions in healthy and clinical populations. We also conduct metascience research to gain insights into research practices within kinesiology and related areas. The lab is located in the Ivor Wynne Centre, which is home to the [Department of Kinesiology](https://www.science.mcmaster.ca/kinesiology/){target="_blank"} within the [Faculty of Science](https://www.science.mcmaster.ca/){target="_blank"} at [McMaster University](https://www.mcmaster.ca/){target="_blank"}.

#### Recent lab news ---------

<font color="#d08770">[July 31 2025].</font> **New Preprint**. Sullivan SR, Buggeln JH, Calalo JA, Ngo TT, Semrau JA, Carter MJ, & Cashaback JGA. Involuntary feedback responses reflect a representation of partner actions. *bioR$\chi$iv*. [https://doi.org/10.1101/2025.07.29.667240](https://doi.org/10.1101/2025.07.29.667240){target="_blank"}

<font color="#d08770">[June 13 2024].</font> **Paper accepted**. Calalo JA, Ngo T, Sullivan S, Strand K, Buggeln JH, Lokesh R, Roth A, Carter MJ, Kurtzer IL, & Cashaback JGA. (Accepted). Online movements reflect ongoing deliberation. *Journal of Neuroscience*. [https://doi.org/10.1523/JNEUROSCI.1913-24.2025](https://doi.org/10.1523/JNEUROSCI.1913-24.2025){target="_blank"}

<font color="#d08770">[May 14 2024].</font> **Paper accepted**. Sullivan SR, Lokesh R, Calalo J, Ngo TT, Buggeln JH, Roth AR, Peters CS, Kurtzer IL, Carter MJ, & Cashaback JGA. (In Press). Indecision under time pressure arises from suboptimal switching behaviour. *Journal of Neurophysiology*. [https://doi.org/10.1152/jn.00563.2024](https://doi.org/10.1152/jn.00563.2024){target="_blank"}