### 2025-08-26 index.qmd
add lab manual link

---
title: ""
pagetitle: "CarterLab | Home"
toc: false
---

![](images/iwc-front.jpg)

The **Action, Cognition, and Metascience Lab** is an interdisciplinary research team interested in how the human brain controls and learns skilled actions in healthy and clinical populations. We also conduct metascience research to gain insights into research practices within kinesiology and related areas. The lab is located in the Ivor Wynne Centre, which is home to the [Department of Kinesiology](https://www.science.mcmaster.ca/kinesiology/){target="_blank"} within the [Faculty of Science](https://www.science.mcmaster.ca/){target="_blank"} at [McMaster University](https://www.mcmaster.ca/){target="_blank"}.

#### Recent lab news ---------

<font color="#d08770">[August 26 2025].</font> **New member**. Welcome to Michael Croteau, who is starting his M.Sc studies. Michael previously completed his undergraduate project and thesis courses in the lab.

<font color="#d08770">[July 31 2025].</font> **New Preprint**. Sullivan SR, Buggeln JH, Calalo JA, Ngo TT, Semrau JA, Carter MJ, & Cashaback JGA. Involuntary feedback responses reflect a representation of partner actions. *bioR$\chi$iv*. [https://doi.org/10.1101/2025.07.29.667240](https://doi.org/10.1101/2025.07.29.667240){target="_blank"}

<font color="#d08770">[June 13 2024].</font> **Paper accepted**. Calalo JA, Ngo T, Sullivan S, Strand K, Buggeln JH, Lokesh R, Roth A, Carter MJ, Kurtzer IL, & Cashaback JGA. (Accepted). Online movements reflect ongoing deliberation. *Journal of Neuroscience*. [https://doi.org/10.1523/JNEUROSCI.1913-24.2025](https://doi.org/10.1523/JNEUROSCI.1913-24.2025){target="_blank"}