### 2025-09-08 teaching.qmd
minor changes and add 2mc3 link

---
title: ""
pagetitle: "CarterLab | Teaching"
toc: false
---

<style>
.course-listing {
  display: flex;
  align-items: center;
  padding: 20px; 
  border: 1px solid #ddd;
  margin-bottom: 20px;
}

.course-img {
  margin-right: 20px;
  width: 175px;
  height: 175px;
  border-radius: 50%;
  object-fit: cover;
}

.course-info {
  display: flex;
  flex-direction: column;
}
</style>

<div class="course-listing">
  <img src="images/logo-2mc3.png" alt="images/logo-2mc3.png" class="course-img">
  <div class="course-info">
  <h4> [Motor Control and Learning](https://carterjmike.github.io/kinesiol-2mc3/) </h4>
  <p> KINESIOL 2MC3 </p>
  <p> Sept-Dec 2025 </p>
  </div>
</div>