### 2025-12-18 index.qmd
add new preprint

---
title: ""
pagetitle: "CarterLab | Home"
toc: false
---

![](images/iwc-front.jpg)

The **Action, Cognition, and Metascience Lab** is an interdisciplinary research team interested in how the human brain controls and learns skilled actions in healthy and clinical populations. We also conduct metascience research to gain insights into research practices within kinesiology and related areas. The lab is located in the Ivor Wynne Centre, which is home to the [Department of Kinesiology](https://www.science.mcmaster.ca/kinesiology/){target="_blank"} within the [Faculty of Science](https://www.science.mcmaster.ca/){target="_blank"} at [McMaster University](https://www.mcmaster.ca/){target="_blank"}.

#### Recent lab news ---------

<font color="#d08770">[December 15 2025].</font> **New Preprint**. Calalo JA, Sullivan SR, Muscara NR, Buggeln JH, Ngo TT, Short MM, Carter MJ, Kurtzer IL, & Cashaback JGA. Deliberation is a controllable process governed by desirability and cognitive effort. *bioR$\chi$iv*. [https://doi.org/10.64898/2025.12.12.693733](https://doi.org/10.64898/2025.12.12.693733 ){target="_blank"}

<font color="#d08770">[September 26 2024].</font> **Paper accepted**. Hussain M, Foglia SD, Park J, Ramdeo KR, Adams FC, Drapeau CC, Bobinski AR, Carter MJ, & Nelson AJ. Ipsilateral contraction increases map area and decreases motor threshold for contralateral hand muscle. *Neuroscience*. [https://doi.org/10.1016/j.neuroscience.2025.09.034](https://doi.org/10.1016/j.neuroscience.2025.09.034){target="_blank"}

<font color="#d08770">[August 26 2025].</font> **New member**. Welcome to Michael Croteau, who is starting his M.Sc studies. Michael previously completed his undergraduate project and thesis courses in the lab.

<font color="#d08770">[July 31 2025].</font> **New Preprint**. Sullivan SR, Buggeln JH, Calalo JA, Ngo TT, Semrau JA, Carter MJ, & Cashaback JGA. Involuntary feedback responses reflect a representation of partner actions.  
*bioR$\chi$iv*. [https://doi.org/10.1101/2025.07.29.667240](https://doi.org/10.1101/2025.07.29.667240){target="_blank"}